Disparities in Climate-Induced Health Outcomes in the Greater Toronto Area.

### 2022-08-08 outputs/paper/paper.qmd
Updates to paper

Urgent action is needed and the science shows that it will make a tangible difference \[\@citeipcc\]. The health impacts from climate change can be mitigated by improving efforts both in societies and health systems in Canada \[\@citeipcc\]. Changes in measures can prevent health impacts and build health systems and facilities that are more resilient to climate change \[\@citeipcc\].

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2012 for December to March", fig.width=9, fig.height=3.2}

level_order <- c("Dec", "Jan", "Feb","Mar")

temp_2002_2022 %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar") & (year >= 2002 & year < 2012)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(. ~ month_name) 
```

Figure 1 shows an increasing pattern for the maximum temperature from year 2002 to 2012 from December to March. Temperatures in March continue to be increasingly higher than any of the other months in this plot. It is noticeable that the temperatures for all months are the highest in 2012. According to NASA, the average temperatures in 2012 was warmer than the mid-20th century baseline and is considered to be an outlying of seasonal extremes that are warmer [@citenasaclimatewarn]. The lowest maximum temperature in January is higher than other years in 2012. The plot shows that the average maximum temperature from December to March has been between -19.2 degrees Celsius in February of 2015 and 24.3 degrees Celsius in March of 2012.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2013 to 2022 for December to March", fig.width=9, fig.height=3.2}

temp_2002_2022 %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar") & (year >= 2012 & year < 2022)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(. ~ month_name)

@@

Figure 2 shows an increasing pattern for the maximum temperature from year 2013 to 2022 from December to March.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2012 for April to July", fig.width=9, fig.height=3.2}

@@

temp_2002_2022 %>% filter((month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul") & (year >= 2002 & year < 2012)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(. ~ month_name) 
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2013 to 2022 for April to July", fig.width=9, fig.height=3.2}

temp_2002_2022 %>% filter((month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul") & (year >= 2012 & year < 2022)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(. ~ month_name) 
```

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2012 for August to November", fig.width=9, fig.height=3.2}

@@

temp_2002_2022 %>% filter((month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov") & (year >= 2002 & year < 2012)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(. ~ month_name)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2013 to 2022 for August to November", fig.width=9, fig.height=3.2}

temp_2002_2022 %>% filter((month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov") & (year >= 2012 & year < 2022)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(. ~ month_name)
```

>>>>

Urgent action is needed and the science shows that it will make a tangible difference [@citeipcc]. The health impacts from climate change can be mitigated by improving efforts both in societies and health systems in Canada [@citeipcc]. Changes in measures can prevent health impacts and build health systems and facilities that are more resilient to climate change [@citeipcc].

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2011 for December to March", fig.width=9, fig.height=3.2}

level_order <- c("Dec", "Jan", "Feb","Mar")
temp_2002_2022$month_name_f = factor(temp_2002_2022$month_name, levels = c("Dec", "Jan", "Feb", "Mar"))

temp_2002_2022 %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar") & (year >= 2002 & year < 2012)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(. ~ month_name_f) 
```

Figure 1 shows an increasing pattern for the maximum temperature from year 2002 to 2011 from December to March. Temperatures in March continue to be increasingly higher than any of the other months in this plot. The lowest maximum temperature in January is higher than other years in 2011. The plot shows that the average maximum temperature from December to March has been between -19.2 degrees Celsius in February of 2015 and 24.3 degrees Celsius in March of 2012.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2012 to 2022 for December to March", fig.width=9, fig.height=3.2}

temp_2002_2022 %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar") & (year >= 2012 & year < 2022)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(. ~ month_name_f)

@@

Figure 2 shows an increasing pattern for the maximum temperature from year 2013 to 2022 from December to March. It is noticeable that the temperatures for all months are the highest in 2012. According to NASA, the average temperatures in 2012 was warmer than the mid-20th century baseline and is considered to be an outlying of seasonal extremes that are warmer [@citenasaclimatewarn].

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2011 for April to July", fig.width=9, fig.height=3.2}

level_order <- c("Apr", "May", "Jun", "Jul")
temp_2002_2022$month_name_f = factor(temp_2002_2022$month_name, levels = c("Apr", "May", "Jun", "Jul"))

temp_2002_2022 %>% filter((month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul") & (year >= 2002 & year < 2012)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(. ~ month_name_f) 
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2012 to 2022 for April to July", fig.width=9, fig.height=3.2}

@@

temp_2002_2022 %>% filter((month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul") & (year >= 2012 & year < 2022)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(. ~ month_name_f) 
```

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2011 for August to November", fig.width=9, fig.height=3.2}

level_order <- c("Aug","Sep", "Oct","Nov")
temp_2002_2022$month_name_f = factor(temp_2002_2022$month_name, levels = c("Aug","Sep", "Oct","Nov"))

temp_2002_2022 %>% filter((month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov") & (year >= 2002 & year < 2012)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(. ~ month_name_f)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2012 to 2022 for August to November", fig.width=9, fig.height=3.2}

@@

temp_2002_2022 %>% filter((month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov") & (year >= 2012 & year < 2022)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(. ~ month_name_f)
```
